---
title: "Hamlet Hapax"
author: "CyberDH"
date: "2015"
output: html_document
---
Words that occur just once are often called _singletons_ (or even _onesies_), but they are more formally called _hapax legomena_ (_hapax_, for short). Here, the method measures lexical variety among play characters and scenes.  

***  
### Global parameters 
```{r}
setwd("~/cyberdh/HILT-demo")
library(knitr)
library(markdown)
library(rmarkdown)
library(ggplot2)
```   

***   
### Hamlet hapax by character
#### Prepare text data, broken into chunks by character
```{r}
playText.s <- scan("hamletCharacters.txt", what="character", sep="\n")
playText.s <- unlist(playText.s)
characterSegments <- grep("^CHARACTER", playText.s)
segmentsEnd <-  length(playText.s)
characterSegments  <-  c(characterSegments , segmentsEnd)
```

***
#### Use loop to calculate word type frequency correlations among segments
```{r}
segmentFreq <- list()
segmentsRaw <- list()
for(i in 1:length(characterSegments)){
  if(i != length(characterSegments)){
    segmentsbyTitle <- playText.s[characterSegments[i]]
    start.segments <- characterSegments[i]+1
    end.segments <- characterSegments[i+1]-1
    segmentLines <- playText.s[start.segments:end.segments]
    segmentWords <- tolower(paste(segmentLines, collapse=" "))
    segmentsStrsplit <- strsplit(segmentWords, "\\W")
    segmentWords <- unlist(segmentsStrsplit)
    segmentWords <- segmentWords[which(segmentWords!="")] 
    segmentTable <- table(segmentWords)
    segmentsRaw[[segmentsbyTitle]] <-  segmentTable
    segmentTable.rel <- 100*(segmentTable/sum(segmentTable))
    segmentFreq[[segmentsbyTitle]] <- segmentTable.rel
  }
}
``` 

***                 
#### Calculate and plot unique word usage   
```{r}
segmentHapax <- sapply(segmentsRaw, function(x) sum(x == 1))
segmentHapax
segment.lengths.m <- do.call(rbind, lapply(segmentsRaw,sum))
hapax.percentage <- segmentHapax / segment.lengths.m
barplot(hapax.percentage, beside=T,col="grey", names.arg = seq(1:length(segmentsRaw)))
```    


***   
### Hamlet hapax by scene
#### Prepare text data, broken into chunks by scene
```{r}
playText.a <- scan("hamlet.txt", what="character", sep="\n")
playText.a <- unlist(playText.a)
scene.positions.v <- grep("^Scene", playText.a)
last.position.v <- length(playText.a)
scene.positions.v <- c(scene.positions.v, last.position.v)
```

***
#### Use loop to calculate word type frequency correlations among scenes
```{r}
scene.raws.l <- list()
scene.freqs.l <- list()

for (i in 1:length(scene.positions.v)) {
  if (i != length(scene.positions.v)) {
    scene.number <- playText.a[scene.positions.v[i]]
    start.positions <- scene.positions.v[i] +1
    end.positions <- scene.positions.v[i+1] -1
    scene.v <- playText.a[start.positions:end.positions]
    scene.words.v <- tolower(paste(scene.v, collapse= " "))
    scene.words.l <- strsplit(scene.words.v, "\\W")
    scene.word.v <- unlist(scene.words.l)
    scene.word.v <- scene.word.v[which(scene.word.v != "")]
    scene.freqs.t <- table(scene.word.v)
    scene.raws.l[[scene.number]] <- scene.freqs.t
    scene.freqs.t.rel <- 100*(scene.freqs.t/sum(scene.freqs.t))
    scene.freqs.l[[scene.number]] <- scene.freqs.t.rel
  }
}
```

***                 
#### Calculate and plot unique word usage   
```{r}
hamletHapax <- sapply(scene.raws.l, function(x) sum(x == 1))
hamletHapax
scene.lengths.m <- do.call(rbind, lapply(scene.raws.l,sum))
scene.percentage <- hamletHapax / scene.lengths.m
barplot(scene.percentage, beside=T,col="grey", names.arg = seq(1:length(scene.raws.l)))
```    


Acknowledgments: Code adapted from Matthew Jockers’ _Text Analysis with R for Students of Literature_.