---
title: "Jane Eyre Dispersion"
author: "CyberDH"
date: "2015"
output: html_document
---

***  
### Global parameters 
#### Set working directory
```{r}
setwd("~/cyberdh/HILT-demo")
```   

#### Include necessary packages for notebook
```{r}
library(knitr)
library(markdown)
library(rmarkdown)
library(ggplot2)
```   

#### Load data (plaintext file), formatted by line breaks  
```{r}
text.v <- scan("janeEyre.txt", what="character", sep="\n")
```  

***   
### Prepare text data
#### Collapse text into one string, leaving spaces " " between words 
```{r}
janeEyre.v <- paste(text.v, collapse="  ")
```   

#### Convert text string to all lower-case letters
```{r}
janeEyre.lower.v <- tolower(janeEyre.v)
```  

#### Split (_strsplit_) big string into separate words
```{r}
janeEyre.words <- strsplit(janeEyre.lower.v, "\\W")
```  

#### Simplify items by reducing the list to a vector
```{r}
janeEyre.word.v <- unlist(janeEyre.words)
```

#### Eliminate blanks left by punctuation
```{r}
not.blanks.v <- which(janeEyre.word.v != "")
janeEyre.word.v <- janeEyre.word.v[not.blanks.v]
```

***   
### Analysis
#### Set all words into number sequence: position = "dramatic time"
```{r}
text.position.v <- seq(1:length(janeEyre.word.v))
```

#### Find dispersion of "love"  
```{r}
love.v <- which(janeEyre.word.v == "love")
love.v
```

***   
#### Create inverse vector of all positions with "NA" that are _not_ love
```{r}
w.count.v <- rep(NA, length(text.position.v))
```

***   
#### Mark all occurrences of "love" = 1
```{r}
w.count.v[love.v] <- 1
```

***   
#### Plot the matrix
```{r}
plot(w.count.v, main="Dispersion Plot of 'love' in Jane Eyre", xlab="Text Location",
      ylab="Love", type="h", ylim=c(0,1), yaxt='n')
```   

***   
#### Find dispersion of "death"  
```{r}
death.v <- which(janeEyre.word.v == "death")
w.count.v <- rep(NA, length(text.position.v))
w.count.v[death.v] <- 1
plot(w.count.v, main="Dispersion Plot of 'death' in Jane Eyre", xlab="Text Location",
      ylab="Death", type="h", ylim=c(0,1), yaxt='n')
```

***   
#### Find dispersion of "little"  
```{r}
little.v <- which(janeEyre.word.v == "little")
w.count.v <- rep(NA, length(text.position.v))
w.count.v[little.v] <- 1
plot(w.count.v, main="Dispersion Plot of 'little' in Jane Eyre", xlab="Text Location",
     ylab="Little", type="h", ylim=c(0,1), yaxt='n')
```

***   
#### Find dispersion of "happy"  
```{r}
happy.v <- which(janeEyre.word.v == "happy")
w.count.v <- rep(NA, length(text.position.v))
w.count.v[happy.v] <- 1
plot(w.count.v, main="Dispersion Plot of 'happy' in Jane Eyre", xlab="Text Location", 
     ylab="Happy", type="h", ylim=c(0,1), yaxt='n')
```

***   
#### Find dispersion of "father"  
```{r}
father.v <- which(janeEyre.word.v == "father")
w.count.v <- rep(NA, length(text.position.v))
w.count.v[father.v] <- 1
plot(w.count.v, main="Dispersion Plot of 'father' in Jane Eyre", xlab="Text Location",
     ylab="Father", type="h", ylim=c(0,1), yaxt='n')
```

Acknowledgments: Code adapted from Matthew Jockers’ Text Analysis with R for Students of Literature.