---
title: "Jane Eyre Top Ten Words"
author: "CyberDH"
date: "2015"
output: html_document
---

***  
### Global parameters 
#### Set working directory
```{r}
setwd("~/cyberdh/HILT-demo")
```

#### Include necessary packages for notebook
```{r}
library(knitr)
library(markdown)
library(rmarkdown)
```

#### Load data (plaintext file), formatted by line breaks  
```{r}
text.v <- scan("janeEyre.txt", what="character", sep="\n")
```

***   
### Prepare text data
#### Collapse text into one string, leaving spaces " " between words 
```{r}
janeEyre.v <- paste(text.v, collapse="  ")
```

#### Convert text string to all lower-case letters
```{r}
janeEyre.lower.v <- tolower(janeEyre.v)
```

#### Split (_strsplit_) big string into separate words
```{r}
janeEyre.words <- strsplit(janeEyre.lower.v, "\\W")
```

#### Simplify items by reducing the list to a vector 
```{r}
janeEyre.word.v <- unlist(janeEyre.words)
```

#### Eliminate blanks left by punctuation; overwrite "janeEyre.word.v"
```{r}
not.blanks.v <- which(janeEyre.word.v != "")
janeEyre.word.v <- janeEyre.word.v[not.blanks.v]
```

***   
### Analysis
#### Number of unique words in the text
```{r}
length(unique(janeEyre.word.v))
```

***    
#### Create a sorted table of “contingency” word types and their corresponding frequencies.
```{r}
janeEyre.freqs.t <- table(janeEyre.word.v)
sorted.janeEyre.freqs.t <- sort(janeEyre.freqs.t, decreasing=TRUE)
```

***   
#### Calculate whether text conforms to Zipf’s Law: _the frequency of any word in a corpus is inversely proportional to its “rank” or position in the overall frequency distribution_.
```{r}
sorted.janeEyre.rel.freqs.t <- 100*(sorted.janeEyre.freqs.t/sum(sorted.janeEyre.freqs.t))
```

***    
#### Plot the result
```{r}
plot(sorted.janeEyre.rel.freqs.t[1:10], type="b", main="Jane Eyre, Entire Novel", xlab="Top Ten Words", 
     ylab="Percentage of Prose", xaxt="n",)
axis(1,1:10, labels=names(sorted.janeEyre.rel.freqs.t[1:10]))
```


Acknowledgments: Code adapted from Matthew Jockers' _Text Analysis with R for Students of Literature_.
